---
title: "scONCO quick-start"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{scONCO quick-start}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

# scONCO — pan-cancer single-cell annotation

`scONCO` annotates a pre-clustered Seurat object with a 4-level pan-cancer
marker hierarchy, adds a v1.1 statistical-confidence layer, and (optionally)
a CNV-aware malignant gate.

## 1. Load the marker database

```{r}
library(scONCO)
markers_df <- load_cancer_marker_db(version = "v1.0", organism = "human")
# Lock to one cancer type (keeps shared TME):
# markers_df <- load_cancer_marker_db(module = "melanoma")
```

## 2. Annotate

```{r}
seu <- run_scONCO(
  seurat_obj       = seu,
  markers_df       = markers_df,
  cluster_col      = "seurat_clusters",
  apply_confidence = TRUE
)
table(seu$scONCO_L4_Abb)
```

## 3. CNV-aware malignant gate (cancer layer)

```{r}
# Supply a precomputed CNV call (e.g. from CopyKAT / scATOMIC), or let the gate
# run CopyKAT itself.
seu <- run_scONCO(seu, markers_df,
                  apply_confidence = TRUE,
                  apply_cnv_gate   = TRUE,
                  cnv_method = "precomputed", cnv_col = "copykat_call")
table(seu$scONCO_malignant_call)
```

## 4. Outputs

`run_scONCO()` writes `scONCO_L1` … `scONCO_L4_Abb`; with `apply_confidence`
it adds `scONCO_L{k}_confidence` / `_status`, `conformal_set`,
`conformal_set_size`; with `apply_cnv_gate` it adds `scONCO_malignant_call`,
`scONCO_malignant_prob`, `scONCO_cnv_score`, `scONCO_L1_gated`,
`scONCO_L4_gated`, `scONCO_gate_corrected`.
